---
title: "Cyberbullying Text Analysis"
format: html
---

```{r}
library(tidyverse)
library(tidymodels)
library(tidytext)

data <- read_csv("https://raw.githubusercontent.com/agmath/agmath.github.io/refs/heads/master/data/classification/cyberbullying_tweets.csv")

data %>%
  count(cyberbullying_type)

data <- data %>%
  distinct()

set.seed(123)
data_split <- initial_split(data, prop = 0.9)

train <- training(data_split)
test <- testing(data_split)
```

```{r}
train %>%
  head()
```

## Tokenization

```{r}
common_words_list <- train %>%
  mutate(tweet_id = row_number()) %>%
  unnest_tokens(word, tweet_text) %>%
  anti_join(stop_words) %>%
  filter(!(word %in% c("http", "https", "t.co", "bully",
                       "bullies", "bullied"))) %>%
  filter(!str_starts(word, "\\d+")) %>%
  count(word) %>%
  arrange(-n) %>%
  filter(n >= 100) %>%
  pull(word)

train %>%
  mutate(tweet_id = row_number()) %>%
  unnest_tokens(word, tweet_text) %>%
  anti_join(stop_words) %>%
  filter(!(word %in% c("http", "https", "t.co", "bully",
                       "bullies", "bullied"))) %>%
  filter(!str_starts(word, "\\d+")) %>%
  filter(word %in% common_words_list) %>%
  distinct() %>%
  slice(1:1e4) %>%
  mutate(
    present = 1
  ) %>%
  pivot_wider(id_cols = c(cyberbullying_type, tweet_id),
              names_from = word,
              values_from = present)

```

```{r}
train %>%
  head()
```


## Data Viz

```{r}
train %>%
  mutate(tweet_id = row_number()) %>%
  unnest_tokens(word, tweet_text) %>%
  anti_join(stop_words) %>%
  filter(!(word %in% c("http", "https", "t.co", "bully",
                       "bullies", "bullied"))) %>%
  filter(!str_starts(word, "\\d+")) %>%
  group_by(cyberbullying_type) %>%
  count(word) %>%
  top_n(15, n) %>%
  ungroup() %>%
  ggplot() + 
  geom_bar(aes(x = reorder_within(word, n, cyberbullying_type), y = n, fill = cyberbullying_type),
           stat = "identity", color = "black",
           show.legend = FALSE) + 
  scale_x_reordered() +
  facet_wrap(~cyberbullying_type, scales = "free") + 
  coord_flip()
```

Compare gender and ethnicity words

```{r}
p <- train %>%
  #filter(str_detect(cyberbullying_type, "gender") | str_detect(cyberbullying_type, "ethnicity"))
  filter((cyberbullying_type == "gender") | (cyberbullying_type == "ethnicity")) %>%
  unnest_tokens(word, tweet_text) %>%
  anti_join(stop_words) %>%
  filter(!(word %in% c("http", "https", "t.co", "bully",
                       "bullies", "bullied"))) %>%
  filter(!str_starts(word, "\\d+")) %>%
  group_by(cyberbullying_type) %>%
  count(word) %>%
  filter(n > 25) %>%
  pivot_wider(names_from = cyberbullying_type, values_from = n) %>% 
  mutate(
    ethnicity = ifelse(is.na(ethnicity), 0, ethnicity),
    gender = ifelse(is.na(gender), 0, gender),
  ) %>%
  ggplot() + 
  geom_text(aes(x = ethnicity, y = gender, label = word)) + 
  geom_abline(linetype = "dashed")

plotly::ggplotly(p)  
```

## Regular Expressions

Extracting Hashtags and Mentions...

```{r}
train %>%
  mutate(
    hashtags = str_extract(tweet_text, "#([A-z]|\\d|-)+"),
    mentions = str_extract(tweet_text, "@([A-z]|\\d|-)+")
  ) %>%
  filter(!is.na(hashtags) | !is.na(mentions))
```

The code above will only extract the first hashtag and first mention for each tweet. We can extract all the hashtags and all the mentions though if we'd like.

```{r}
train %>%
  mutate(
    hashtags = str_extract_all(tweet_text, "#([A-z]|\\d|-)+"),
    mentions = str_extract_all(tweet_text, "@([A-z]|\\d|-)+")
  ) %>%
  filter((lengths(hashtags) > 0) | (lengths(mentions) > 0)) %>%
  unnest(mentions) %>%
  unnest(hashtags)
```

Notice that every tweet now is repeated once for every combination of hashtag and mention. For example, the tweet `\@TheDemocrats and the "I don't care" \@gop Against our better judgement...` appears six times in the data frame above. This is because there were two hashtags and three mentions in the tweet ($2\times 3 = 6$). This format might be okay, but we could obtain different formats if we like -- for example, one row per tweet.

```{r}
train %>%
  mutate(
    hashtags = map_chr(str_extract_all(tweet_text, "#([A-z]|\\d|-)+"), ~ paste(.x, collapse = ", ")),
    mentions = map_chr(str_extract_all(tweet_text, "@([A-z]|\\d|-)+"), ~ paste(.x, collapse = ", "))
  ) %>%
  filter((hashtags != "") | (mentions != ""))
```

In the interest of full transparency, I wasn't going to approach constructing this particular data frame in this way. My code was going to be much lengthier and messier. ChatGPT was quite helpful in simplifying the code.

We could now do things like count the number of hashtags and the number of mentions.

```{r}
train %>%
  mutate(
    hashtags = map_chr(str_extract_all(tweet_text, "#([A-z]|\\d|-)+"), ~ paste(.x, collapse = ", ")),
    mentions = map_chr(str_extract_all(tweet_text, "@([A-z]|\\d|-)+"), ~ paste(.x, collapse = ", "))
  ) %>%
  filter((hashtags != "") | (mentions != "")) %>%
  mutate(
    hashtag_count = str_count(hashtags, ",") + 1,
    mention_count = str_count(mentions, ",") + 1,
  )
```



